---
title: "Crime Data Explonatory Analysis Project"
format: html
editor: visual
---

## Data Set
We are thinking about dealing with crime dataset. Here is the link of the data we are planning to used: https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8

## Data set Description
This dataset reflects incidents of crime in the City of Los Angeles dating back to 2020. There are 521k rows (each row is a crime incident) and 28 columns (Variables).

1. *DR_NO* - Division of records number: Official file number made up of a 2 digit year, area ID, and 5 digits.

2. *Date Rptd* - Reported date of the crime

3. *Date OCC* - Occured date of the crime

4. *TIME OCC* - occured time of the crime in 24 hour military time.

5. *AREA* - The Los Angeles Police Department has 21 community police Stations referred to as Geographic Areas within the department. These Geographic Areas are sequentially numbered from 1-21. 

6. *AREA NAME* - Name designation of the area.

7. *Rpt Dist No* - A four digit code that represents a sub area with a geographic Area.

8. *Part 1-2* - no discription

9. *Crm cd* - indicates the crime committed (same as the *Crm cd 1*)

10. *Crm cd Desc* - Defines the crime Code Provided.

11. *Mocodes* - Activities associated with the suspect in commission of the crime.

12. *Vict Age* - Age of the Victim.

13. *Vict Sex* - Sex of the Victim (*F* - Female/ *M* - Male, *X*- Unknown)

14. *Vict Descent* - Descent Code of the Victim (*A*-other Asian, *B*-Black, *C*-Chinese, etc.)

15. *Premis Cd* - The type of structure, vehicle, or location where the crime took place.


## Potential Topics to Explore Using the Data Set

We will explore following facts using the data.

1. Which area has the highest number of crimes among the 21 Police Stations.

2. Create a table for area and the sub area to identify the most awful sub area.

3. Use *Crm cd* and *Crm cd Desc* data to explore Which is the most frequent crime type.

4. Present a Side by side box plot for the ages with most frequent 5 types of  crime in order to identify whether there is a targeted age for particular crime. 

5. Bar plot of crime type (crm cd) filled with victim sex, descent in order to identify whether there is a targeted sex, descent for particular crime.


## Explonatory Analysis



```{r}

```



```{r}

```


