---
title: "Exploratory Analysis of a Crime Data"
format: html
editor: visual
---

```{r,include=FALSE}
library(tidyverse)
```

## Data Set

Here is the link of the data we used for our analysis: https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8

## Data set Description

This dataset reflects incidents of crime in the City of Los Angeles dating back to 2020. There are 521k rows (each row is a crime incident) and 28 columns (Variables).

1.  *DR_NO* - Division of records number: Official file number made up of a 2 digit year, area ID, and 5 digits.

2.  *Date Rptd* - Reported date of the crime

3.  *Date OCC* - Occured date of the crime

4.  *TIME OCC* - occured time of the crime in 24 hour military time.

5.  *AREA* - The Los Angeles Police Department has 21 community police Stations referred to as Geographic Areas within the department. These Geographic Areas are sequentially numbered from 1-21.

6.  *AREA NAME* - Name designation of the area.

7.  *Rpt Dist No* - A four digit code that represents a sub area with a geographic Area.

8.  *Part 1-2* - no discription

9.  *Crm cd* - indicates the crime committed (same as the *Crm cd 1*)

10. *Crm cd Desc* - Defines the crime Code Provided.

11. *Mocodes* - Activities associated with the suspect in commission of the crime.

12. *Vict Age* - Age of the Victim.

13. *Vict Sex* - Sex of the Victim (*F* - Female/ *M* - Male, *X*- Unknown)

14. *Vict Descent* - Descent Code of the Victim (*A*-other Asian, *B*-Black, *C*-Chinese, etc.)

15. *Premis Cd* - The type of structure, vehicle, or location where the crime took place.

16. *Premis Desc* Defines the Premis Cd provided.

17. *Weapon Used Cd* The type of weapon used in crime given as a code.

18. *Weapon Desc* Defines the code of the weapon used in crime.

19. *Status* Provides the status of the case in code.

20. *Status Desc* Description of the status code provided.

21. *Crm Cd 1* It indicates the crime committed where Crime Code 1 represents the most serious crime.

22. *Crm Cd 2* It also indicates the crime code but it is less serious than Crm Cd 1.

23. *Crm Cd 3* This crime code is for less serious crime than represented by Crm Cd 1 and Crm Cd 2.

24. *Crm Cd 4* This crime code is for the least serious crime.

25. *LOCATION* Street address of crime incident rounded to nearest hundred blocks.

26. *Cross Street* Cross street of the rounded address.

27. *LAT* Latitude.

28. *LON* Longitude.

### Loading data

```{r}
data<-read.csv("Subset_data.csv")
```


## Method and analsis

```{r}
length(unique(data$Crm.Cd.Desc))
```


We found that there are `r length(unique(data$Crm.Cd.Desc))` unique crime descriptions but most of the crimes have common characteristics even though the Crime description is different to each other. Therefore, for the convenience of the analysis the variable (`Crm.Cd.Desc`) representing the type of crimes were coded into general categories. The re-categorization was like if the crime contained the word Assualt then crime was categorized as `ASSUALT`, if the crime contained the word Burglary or Stolen or Theft then the crime was coded as `BURGLARY`, if the crime contained the word Criminal or Weapon or Shots or Threat or Kidnap then the crime was coded as `Extreme crimes`, if the crime contained the word Sex or Rape then the crime was coded as `SEXUAL MISCONDUCT`, some crimes were coded as their names like Violation,Robbery, and Trespassing, and rest of the crimes are coded as `OTHER CRIMES`. The missing values of variable `Vict.Sex` was coded as unknown and replace into the same variable.

#### Codes for recoding crime type and victim sex

```{r}
for(i in 1:length(data$Crm.Cd.Desc)){
if(str_detect(data$Crm.Cd.Desc[i],"ASSAULT")==TRUE & 
   str_detect(data$Crm.Cd.Desc[i],"SEXUAL")!=TRUE){
  data$crime[i]<-"ASSAULT"
} else if(str_detect(data$Crm.Cd.Desc[i],"VIOLATION")==TRUE){
  data$crime[i]<-"VIOLATION"
}  else if(str_detect(data$Crm.Cd.Desc[i],"BURGLARY")==TRUE|
           str_detect(data$Crm.Cd.Desc[i],"STOLEN")==TRUE|
            str_detect(data$Crm.Cd.Desc[i],"THEFT")==TRUE){
  data$crime[i]<-"BURGLARY"
} else if(str_detect(data$Crm.Cd.Desc[i],"SEX")==TRUE|str_detect(data$Crm.Cd.Desc[i],"RAPE")==TRUE){
  data$crime[i]<-"SEXUAL MISCONDUCT"
} else if(str_detect(data$Crm.Cd.Desc[i],"VANDALISM")==TRUE){
  data$crime[i]<-"VANDALISM"
} else if(str_detect(data$Crm.Cd.Desc[i],"CRIMINAL")==TRUE|
          str_detect(data$Crm.Cd.Desc[i],"WEAPON")==TRUE|
          str_detect(data$Crm.Cd.Desc[i],"SHOTS")==TRUE|
          str_detect(data$Crm.Cd.Desc[i],"THREAT")==TRUE|
          str_detect(data$Crm.Cd.Desc[i],"KIDNAP")==TRUE){
  data$crime[i]<-"EXTREME CRIMES"
} else if(str_detect(data$Crm.Cd.Desc[i],"ROBBERY")==TRUE){
  data$crime[i]<-"ROBBERY"
} else if(str_detect(data$Crm.Cd.Desc[i],"TRESPASSING")==TRUE){
  data$crime[i]<-"TRESPASSING"
}else data$crime[i]<-"OTHER CRIMES"
}
data$Vict.Sex<-ifelse(data$Vict.Sex=="M","Male",ifelse(data$Vict.Sex=="F","Female","Unknown"))
```




### Victim's descent and crime
```{r}
knitr::kable(data %>% 
  select(Vict.Descent) %>% 
  count(Vict.Descent) %>% 
  rename("Victim descent"="Vict.Descent"))

```

We focused about the descent of the Victims and identified that mainly victims belong to 5 types of descent and rest of are minor or unknown. So we re-categorized the variable `Vict.Descent` into Asian, Black, Hispanic/Latin/Mexican, Other, White, and Minor or Unknown.

```{r}
data$Vict.Descent<-ifelse(data$Vict.Descent=="A","Other Asian",
                          ifelse(data$Vict.Descent=="B","Black",
                            ifelse(data$Vict.Descent=="H","Hispanic/Latin/Mexican",
                              ifelse(data$Vict.Descent=="O","Other",
                                ifelse(data$Vict.Descent == "W","White","Minor or Unknown")
                                        ))))
```


```{r}
data %>%
  group_by(crime,Vict.Descent) %>% 
  count() %>% 
  ungroup() %>% 
      ggplot()+
        aes(x=reorder(crime,n),y=n,fill=Vict.Descent)+
        geom_col(position = "dodge")+
        coord_flip()+
          scale_fill_manual(values = c("#242424", "#00F5FF","#FF3E96" , "#00FF7F", "#FF8247", "#FFF68F"))+
          labs(x="List of Crimes",title="Visual representation between victim's descent and crime type",fill="Victim's Descent")+
          theme_bw()
```
We thought it is better to neglect the minor & unknown category, since we can't get an idea about the descent. So, Victims related to Hispanic/Latin/Mexican have faced every type of crime more except for the burglary, where the white people have faced it more. We also see that assault is faced by a considerable number of black people.


### Victim's age and crime

Now we will consider victims' age. At first we are going to observe whether there are any unusual observations in the data set.

```{r}
hist(data$Vict.Age,xlab ="Victim's Age",main="Histogram of victim's age",breaks =15)
```

From the histogram, we can clearly see that there are a good number of observations recorded under 0 (years) age. We interpreted 0 as the missing records. So, we decided to replace them with the median of the victims' age after deleting the value 0 since the histogram is skewed.

```{r}
data$Vict.Age <- ifelse(data$Vict.Age == "0", median(data$Vict.Age[data$Vict.Age!=0]),data$Vict.Age)
```


```{r}
ggplot(data=data)+
  aes(x=crime,y=Vict.Age)+
  geom_boxplot()+
  labs(x="List of crimes",
       y="Age",title ="Age variation according to the  crimes")+
  coord_flip()
  
```

This Box plot gives us a nice interpretation. When we focused about the IQR we can roughly say majority of victims are in between 25 and 50 for all the crimes except the sexual misconduct. Criminals have targeted a lower age ranges while committing crimes related to the sex.


### Area and reported frequency of crime

The Los Angeles Police Department has 21 community police Stations referred to as Geographic Areas within the department. Now we are trying to come up with an answer, which geographic area has the highest number of crimes recorded.

```{r}
ggplot(data , aes(x = AREA.NAME)) + 
  geom_bar(fill="#FF8247") + 
  geom_text(stat='count', aes(label=..count..), hjust=-0.07,vjust= 0.5)+ coord_flip()+
  theme_bw()+
  labs(x="Area name",title="Frequency of crime occurrence in the areas")
```

According to this bar chart it is obvious that the 77th street has the highest number of crimes occurred.Among the 21 geographic areas, 100+ crimes have been reported over areas named Southwest, Southeast, Pacific, N Hollywood etc. So be careful in those area of Los Angeles, the statistics are not good.

### Occurrence of crime in different areas

This set of Bar plots give an idea regrading what type of crimes frequently happened in Particular area. Since Burglary and Assault has a significant occurrence, we are reporting them separately.

```{r}
data %>% 
  filter(crime=="BURGLARY") %>%
  group_by(AREA.NAME) %>% 
   tally() %>% 
  mutate(rel=prop.table(n)) %>% 
ggplot(aes(x = AREA.NAME,y=rel)) + 
  geom_col(fill="peachpuff3") + coord_flip()+
  labs(x="Area name",y="Relative frequency",title="Proportion for the occurence of crime BURGLARY in different areas")+
  theme_bw()
```

We observed that the occurrence is proportionally highest in area Pacific and lowest Southeast. There are also good proportion of burglary in N Hollywood, Central, and West LA.

```{r}
data %>% 
  filter(crime=="ASSAULT") %>%
  group_by(AREA.NAME) %>% 
   tally() %>% 
  mutate(rel=prop.table(n)) %>% 
ggplot(aes(x = AREA.NAME,y=rel)) + 
  geom_col(fill="#FF3E96") + coord_flip()+
  labs(x="Area name",y="Relative frequency",title="Proportion for the occurence of crime ASSAULT in different areas")
```
We observed that the occurrence of ASSAULT is proportionally highest in 77th street and lowest Devonshire. There are also good proportion of assault in Hollywood, Central, Southeast, Olympic and Southwest..



### Crime and Victim Sex

Now we will be exploring which sex group is experiencing which type of crime more or less. To explore that we are filling the occurrence of crimes with respect to the sex of the victim.

```{r}
data %>%
  group_by(crime,Vict.Sex) %>% 
  count() %>% 
  ungroup() %>% 
      ggplot()+
        aes(x=reorder(crime,n),y=n,fill=Vict.Sex)+
        geom_col(position = "dodge")+
        coord_flip()+
          scale_fill_manual(values = c("#AE4371", "lightblue", "lightyellow"))+
          labs(x="List of Crimes",title="Visual representation of which Sex group is victim of which crime")+
          theme_bw()
```

The graph gives us a view that burglary, assault, vandalism, robbery and extreme crimes are happening with the males more compared to females and people with un-identified sex in the data. The graph also demonstrates that females are more victim of crime like violation and sexual misconduct compared to males. It is also interesting to report that the crimes other than the specified ones are occurring both equally with males and females.


## About premises of crime

Now we will move to explore the premises of different crimes we defined for the convenience of our analysis. We are reporting the premises which appeared more than 3% of the times for the crimes.

```{r}
knitr::kable(data %>% 
  filter(crime=="BURGLARY") %>% 
  count(Premis.Desc) %>% 
  mutate(prop=prop.table(n)) %>% 
  filter(prop>0.030) %>% 
  select(Premis.Desc) %>% 
    rename("Premis of Burglary"="Premis.Desc"))
```

We observe that driveway, garage, multi-unit or single family dwelling, street, parking lot are the frequently occurred premises for the crime burglary.

```{r}
knitr::kable(data %>% 
  filter(crime=="ASSAULT") %>% 
  count(Premis.Desc) %>% 
  mutate(prop=prop.table(n)) %>% 
  filter(prop>=0.03) %>% 
  select(Premis.Desc) %>% 
    rename("Premis of Assault"="Premis.Desc"))
```

We also observe that assault is being reported frequently taking place in street, sidewalk, parking lot, and single family or multi-unit dwelling.

```{r}
knitr::kable(data %>% 
  filter(crime=="SEXUAL MISCONDUCT") %>% 
  count(Premis.Desc) %>% 
  mutate(prop=prop.table(n)) %>% 
  filter(prop>=0.03) %>% 
  select(Premis.Desc) %>% 
    rename("Premis of Sexual Misconduct"="Premis.Desc"))
```

The frequent premises of sexual misconduct is similar to the premises that were reported for burglary and assault but the interesting is that we found that police facility is a frequently occurred premise for the crime sexual misconduct.

```{r}
knitr::kable(data %>% 
  filter(crime=="EXTREME CRIMES") %>% 
  count(Premis.Desc) %>% 
  mutate(prop=prop.table(n)) %>% 
  filter(prop>=0.03) %>% 
  select(Premis.Desc) %>% 
    rename("Premis of Extreme Crimes"="Premis.Desc"))
```

As expected from the frequently occurred premises of previous crimes, the extreme crimes are also frequently happening in the street, parking lot, sidewalk, and dwelling's. But here alley is a new premise we observed in case of extreme crimes.

## About weapons used in crime

In this part we explored the question that whether for committing a particular type of crime commit-er has preference over the weapons based on the victim's sex. We reported the weapons used more than 10 times for committing a particular crime in the data-set.

##### Assault

In line with that the first crime we explored is Assault.

```{r}
weapon<-data %>% 
  filter(crime=="ASSAULT") %>% 
  count(Weapon.Desc) %>% 
  filter(n>=10) %>% 
  select(Weapon.Desc)
weapon<-weapon$Weapon.Desc  
weapon2<-data %>% 
  filter(crime=="ASSAULT",Weapon.Desc %in% weapon) %>% 
    select(Weapon.Desc,Vict.Sex)
knitr::kable(table(weapon2$Weapon.Desc,weapon2$Vict.Sex))
```

From the contingency table, it is observed that hand gun, semi-automated pistol are more used in case of Male for committing assault compared to in case of female. And bodily force is used more in case of female compared to male.

##### Sexual Misconduct

In this part we explored about the weapons used in terms of victim's sex for the crime sexual misconduct.

```{r}
weapon<-data %>% 
  filter(crime=="SEXUAL MISCONDUCT") %>% 
  count(Weapon.Desc) %>% 
  filter(n>=10) %>% 
  select(Weapon.Desc)
weapon<-weapon$Weapon.Desc  
weapon2<-data %>% 
  filter(crime=="SEXUAL MISCONDUCT",Weapon.Desc %in% weapon) %>% 
  select(Weapon.Desc,Vict.Sex)
knitr::kable(table(weapon2$Weapon.Desc,weapon2$Vict.Sex))
```

We observe that in case of committing sexual miscoduct bodily force is mostly used in case of female compared to male's.

##### Robbery

Let's look for robbery now.

```{r}
weapon<-data %>% 
  filter(crime=="ROBBERY") %>% 
  count(Weapon.Desc) %>% 
  filter(n>=10) %>% 
  select(Weapon.Desc)
weapon<-weapon$Weapon.Desc  
weapon2<-data %>% 
  filter(crime=="ROBBERY",Weapon.Desc %in% weapon) %>% 
  select(Weapon.Desc,Vict.Sex)
knitr::kable(table(weapon2$Weapon.Desc,weapon2$Vict.Sex))
```

As cliche as it sounds, the commit-er has preference for hand gun, semi-automated pistol, and bodily force more for male's compared to female's in case of committing robbery.

##### Extreme crimes

Though not that much extreme crimes in the dataset still let's look at the weapons preference.

```{r}
weapon<-data %>% 
  filter(crime=="EXTREME CRIMES") %>% 
  count(Weapon.Desc) %>% 
  filter(n>=10) %>% 
  select(Weapon.Desc)
weapon<-weapon$Weapon.Desc  
weapon2<-data %>% 
  filter(crime=="EXTREME CRIMES",Weapon.Desc %in% weapon) %>% 
  select(Weapon.Desc,Vict.Sex)
knitr::kable(table(weapon2$Weapon.Desc,weapon2$Vict.Sex))
```

Here you go we have found verbal threats are given frequently in case of committing extreme crimes in the data-set with hand gun and they are more used in case of males compared to females.


## About crime and case status

This explores which type of cases are getting resolved mostly and which type are not. To do this we re-categorized the variable indicating case status into "Not Resolved" if the investigation is continued otherwise coded as "Resolved".

```{r}
data %>% 
  mutate(status=ifelse(Status.Desc=="Invest Cont","Not Resolved","Resolved")) %>% 
  select(crime,status) %>% 
  group_by(crime,status) %>%
  tally() %>% 
  mutate(rel=prop.table(n)) %>% 
  ggplot()+
  aes(x=reorder(crime,n),y=rel,fill=status) %>% 
  geom_col(position = "dodge")+
  coord_flip()+
  scale_fill_manual(values = c("peachpuff3", "lightskyblue"))+
  labs(x="List of crimes",y="Relative frequency",title="Status of the crimes",fill="Status")
```

The graph demonstrates that the crime violation is getting resolved more than any any other crimes in the data-set. Trespassing, burglary, robbery, and sexual misconduct has a fairly large difference between proportion of resolved and not resolved cases. Assault and extreme crime has shorter difference between proportion resolved and not resolved.




