---
title: "Crime Data Explonatory Analysis Project"
format: html
editor: visual
---

```{r,include=FALSE}
library(tidyverse)
```

## Data Set

We are thinking about dealing with crime dataset. Here is the link of the data we are planning to used: https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8

## Data set Description

This dataset reflects incidents of crime in the City of Los Angeles dating back to 2020. There are 521k rows (each row is a crime incident) and 28 columns (Variables).

1.  *DR_NO* - Division of records number: Official file number made up of a 2 digit year, area ID, and 5 digits.

2.  *Date Rptd* - Reported date of the crime

3.  *Date OCC* - Occured date of the crime

4.  *TIME OCC* - occured time of the crime in 24 hour military time.

5.  *AREA* - The Los Angeles Police Department has 21 community police Stations referred to as Geographic Areas within the department. These Geographic Areas are sequentially numbered from 1-21.

6.  *AREA NAME* - Name designation of the area.

7.  *Rpt Dist No* - A four digit code that represents a sub area with a geographic Area.

8.  *Part 1-2* - no discription

9.  *Crm cd* - indicates the crime committed (same as the *Crm cd 1*)

10. *Crm cd Desc* - Defines the crime Code Provided.

11. *Mocodes* - Activities associated with the suspect in commission of the crime.

12. *Vict Age* - Age of the Victim.

13. *Vict Sex* - Sex of the Victim (*F* - Female/ *M* - Male, *X*- Unknown)

14. *Vict Descent* - Descent Code of the Victim (*A*-other Asian, *B*-Black, *C*-Chinese, etc.)

15. *Premis Cd* - The type of structure, vehicle, or location where the crime took place.

16. *Premis Desc* Defines the Premis Cd provided.

17. *Weapon Used Cd* The type of weapon used in crime given as a code.

18. *Weapon Desc* Defines the code of the weapon used in crime.

19. *Status* Provides the status of the case in code.

20. *Status Desc* Description of the status code provided.

21. *Crm Cd 1* It indicates the crime committed where Crime Code 1 represents the most serious crime.

22. *Crm Cd 2* It also indicates the crime code but it is less serious than Crm Cd 1.

23. *Crm Cd 3* This crime code is for less serious crime than represented by Crm Cd 1 and Crm Cd 2.

24. *Crm Cd 4* This crime code is for the least serious crime.

25. *LOCATION* Street address of crime incident rounded to nearest hundred blocks.

26. *Cross Street* Cross street of the rounded address.

27. *LAT* Latitude.

28. *LON* Longitude.

## Explonatory Analysis

##### Here which sex group are the victim of which type of crimes.

```{r}
data<-read.csv("Subset_data.csv")
```

```{r}
for(i in 1:length(data$Crm.Cd.Desc)){
if(str_detect(data$Crm.Cd.Desc[i],"ASSAULT")==TRUE & 
   str_detect(data$Crm.Cd.Desc[i],"SEXUAL")!=TRUE){
  data$crime[i]<-"ASSAULT"
} else if(str_detect(data$Crm.Cd.Desc[i],"VIOLATION")==TRUE){
  data$crime[i]<-"VIOLATION"
}  else if(str_detect(data$Crm.Cd.Desc[i],"BURGLARY")==TRUE|
           str_detect(data$Crm.Cd.Desc[i],"STOLEN")==TRUE|
            str_detect(data$Crm.Cd.Desc[i],"THEFT")==TRUE){
  data$crime[i]<-"BURGLARY"
} else if(str_detect(data$Crm.Cd.Desc[i],"SEX")==TRUE|str_detect(data$Crm.Cd.Desc[i],"RAPE")==TRUE){
  data$crime[i]<-"SEXUAL MISCONDUCT"
} else if(str_detect(data$Crm.Cd.Desc[i],"VANDALISM")==TRUE){
  data$crime[i]<-"VANDALISM"
} else if(str_detect(data$Crm.Cd.Desc[i],"CRIMINAL")==TRUE|
          str_detect(data$Crm.Cd.Desc[i],"WEAPON")==TRUE|
          str_detect(data$Crm.Cd.Desc[i],"SHOTS")==TRUE|
          str_detect(data$Crm.Cd.Desc[i],"THREAT")==TRUE|
          str_detect(data$Crm.Cd.Desc[i],"KIDNAP")==TRUE){
  data$crime[i]<-"CWSTK"
} else if(str_detect(data$Crm.Cd.Desc[i],"ROBBERY")==TRUE){
  data$crime[i]<-"ROBBERY"
} else if(str_detect(data$Crm.Cd.Desc[i],"TRESPASSING")==TRUE){
  data$crime[i]<-"TRESPASSING"
}else data$crime[i]<-"OTHER CRIMES"
}

data$Sex<-ifelse(data$Vict.Sex=="M","Male",ifelse(data$Vict.Sex=="F","Female","Unknown"))
```

```{r}
data %>%
  group_by(crime,Sex) %>% 
  count() %>% 
  ungroup() %>% 
      ggplot()+
        aes(x=reorder(crime,n),y=n,fill=Sex)+
        geom_col(position = "dodge")+
        coord_flip()+
          scale_fill_manual(values = c("blue", "black", "cyan"))+
          labs(x="List of Crimes")+
          theme_bw()
```
```{r}
#add new column that cuts "Victim's age" into categories
data$cat_age <- cut(data$Vict.Age,
              breaks=c(10,30,50,70,100),
              labels=c('child', 'Young', 'Adult', 'Senoir'))

data %>%
  group_by(crime,cat_age) %>% 
  count() %>% 
  ungroup() %>% 
  ggplot()+
  aes(x=reorder(crime,n),y=n,fill=cat_age)+
  geom_col(position = "dodge")+
  coord_flip()+
  scale_fill_manual(values = c("blue","red", "black", "cyan"))+
  labs(x="List of Crimes")+
  theme_bw()


```

##### Here the premises of some crimes are explored given that the premise appeared more than 3% of the times.

```{r}
knitr::kable(data %>% 
  filter(crime=="BURGLARY") %>% 
  count(Premis.Desc) %>% 
  mutate(prop=prop.table(n)) %>% 
  filter(prop>0.030) %>% 
  select(Premis.Desc) %>% 
    rename("Premis of Burglary"="Premis.Desc"))
```

```{r}
knitr::kable(data %>% 
  filter(crime=="ASSAULT") %>% 
  count(Premis.Desc) %>% 
  mutate(prop=prop.table(n)) %>% 
  filter(prop>=0.03) %>% 
  select(Premis.Desc) %>% 
    rename("Premis of Assault"="Premis.Desc"))
```

```{r}
knitr::kable(data %>% 
  filter(crime=="SEXUAL MISCONDUCT") %>% 
  count(Premis.Desc) %>% 
  mutate(prop=prop.table(n)) %>% 
  filter(prop>=0.03) %>% 
  select(Premis.Desc) %>% 
    rename("Premis of Sexual Misconduct"="Premis.Desc"))
```

```{r}
knitr::kable(data %>% 
  filter(crime=="CWSTK") %>% 
  count(Premis.Desc) %>% 
  mutate(prop=prop.table(n)) %>% 
  filter(prop>=0.03) %>% 
  select(Premis.Desc) %>% 
    rename("Premis of CWSTK"="Premis.Desc"))
```
